\begin{thebibliography}{9}

\bibitem{hammerstrom2007} Hammerstrom, D. J., et al., ``Pacific Northwest Gridwise Testbed: Part I - Olympic Peninsula Demonstration.'' PNNL Report 17167, Pacific Northwest National Laboratory, Richland WA, 2007. 

\bibitem{widergren2014} Widergren, S. E., et al., ``AEP Ohio gridSMART,'' PNNL Report 23192, Pacific Northwest National Laboratory, Richland WA, 2014. 

\bibitem{schweppe1978} Schweppe, F. C., ``Power Systems 2000: hierarchical control strategies,'' IEEE Spectrum 15:7, July 1978. DOI: 10.1109/MSPEC.1978.6367813

\bibitem{huberman1991} Huberman, B., S. H. Clearwater, and T. Hogg, ``Cooperative Solution of Constraint Satisfaction Problems,'' JSTOR 254:5035, pp. 1181-1183, 1991. 

\bibitem{kok2013} Kok, K., ``The PowerMatcher Smart Coordination for the Smart Electricity Grid,'' TNO, The Netherlands, 2013.

\bibitem{sepa2019} SEPA, ``Transactive Energy: Real-World Applications for the Modern Grid,'' \textit{Smart Electric Power Alliance}, April 2019. URL:  https://sepapower.org/resource/transactive-energy-real-world-applications-for-the-grid/

\bibitem{chassin2015} Chassin, D.P., Stoustrup, J., Agathoklis, P., Djilali, N.,  ``A new thermostat for real-time price demand response: Cost, comfort and energy impacts of discrete-time control without deadband,'' \textit{Applied Energy} 115, 816--825, 2015.

\bibitem{DOE-G413-2009} US Department of Energy, Guide 413.3-4, ``US Department of Energy Technology Readiness Assessment Guide,'' 10-12-2009. URL: www.directives.doe.gov/directives-documents/400-series/0413.3-EGuide-04

\bibitem{hu2019} Hu, W-L., C. Rivetta, E. MacDonald, D.P. Chassin, ``Modeling of Operator Performance for Human-in-the-loop Power Systems,'' In \textit{International Conference on Human-Computer Interaction}, HCII Engineering Psychology and Cognitive Ergonomics, pp 39-54, 2019.

\end{thebibliography}

